Comparison of Fast Imaging Architectures for Large, Multi-scale, Hierarchical Aperture Arrays

### 2024-06-14 ms.tex
Update on Overleaf.

\begin{abstract}
Two major science drivers of modern radio astronomy, namely, explosive astrophysical transient phenomena and observational cosmology, are pushing aperture arrays towards large-$N$, low-cost element design consisting of multiple spatial scales. Multi-scale aperture arrays require hybrid combinations of data processing architectures-- pre-correlation beamformer, generic version of FFT-based direct imager, post-correlation beamformer, and post-correlation FFT imager-- operating on different spatial scales to obtain optimal performance in imaging the entire field of view. We use the metric of computational cost density defined over the dimensions of discovery space, namely, field of view, angular resolution, polarisation, frequency, and time to determine the most efficient hybrid architecture for aperture arrays spanning multiple spatial scales like the element size, stations (a collection of elements), and the array (a collection of stations) with a range of number of elements per station and number of stations. We use nominal parameters of specific upcoming and planned arrays-- Square Kilometre Array (SKA) at low frequencies (\texttt{SKA-low}), \texttt{SKA-low-core}, a long baseline extension to SKA-low (\texttt{LAMBDA-I}), compact all-sky phased array (CASPA), and a lunar array (\texttt{FarView})-- to determine the most optimal architectural hierarchy from a computational standpoint, and provide a guide for designing hybrid architectures for multi-scale aperture arrays. For large-$N$, dense-packed layouts, a FFT-based direct imager is most efficient for most cadence intervals, and for other layouts the best architecture depends strongly on the cadence interval.
\end{abstract}

>>>>

\begin{abstract}
Two major science drivers of modern radio astronomy, namely, explosive astrophysical transient phenomena and observational cosmology, are pushing aperture arrays towards large-$N$, low-cost element design consisting of multiple spatial scales. Multi-scale aperture arrays require hybrid combinations of data processing architectures-- pre-correlation beamformer, generic version of FFT-based direct imager, post-correlation beamformer, and post-correlation FFT imager-- operating on different spatial scales to obtain optimal performance in imaging the entire field of view. We use the metric of computational cost density defined over the dimensions of discovery space, namely, field of view, angular resolution, polarisation, frequency, and time to determine the most efficient hybrid architecture for aperture arrays spanning multiple spatial scales like the element size, stations (a collection of elements), and the array (a collection of stations) with a range of number of elements per station and number of stations. We use nominal parameters of specific upcoming and planned arrays-- Square Kilometre Array (SKA) at low frequencies (\texttt{SKA-low}), \texttt{SKA-low-core}, a long baseline extension to SKA-low (\texttt{LAMBDA-I}), compact all-sky phased array (CASPA), and a lunar array (\texttt{FarView})-- to determine the most optimal architectural hierarchy from a computational standpoint, and provide a guide for designing hybrid architectures for multi-scale aperture arrays. For large-$N$, dense-packed layouts, a FFT-based direct imager is most efficient for most cadence intervals, and for other layouts that have relatively lesser number of elements or greater sparsity in distribution, the best architecture is more sensitive to the cadence interval.
\end{abstract}
